---
title: "2020-12-17 Exploring Michale's Data"
output: html_notebook
---

## Objective 

* Take a look at what kind of data MW actually sent us, is it monthly or annual data? 
* Define the functions that can be used to to calculate the trend line 
  * linear regression
  * smooth regression 
  * 20 EOC - historical mean
  
This will most likely be with cmip5

# CMIP 5 

## Taking a look at the MW's data! 

```{r}
library(ncdf4)
library(magrittr)
library(ggplot2)
library(raster)
library(assertthat)
```

```{r}
INPUT <- here::here('inputs', 'metric')
```


```{r}
nc1 <- nc_open(file.path(INPUT, 'TN_CESM1-CAM5_historical_rcp85_r10i1p1_192001-210012.nc'))
TNx_d <- ncvar_get(nc1, 'TNx')
t <- ncvar_get(nc1, 'time')
ncatt_get(nc1, 'time')
summary(unlist(TNx_d))
1800 + 120
1800 + 300
```

So it looks like the TN file returns TNx values, min daily temperature with the variable name 'TNx'.

```{r}
nc2 <- nc_open(file.path(INPUT, 'TX_CanESM2_historical_rcp85_r10i1p1_19500101-21001231.nc'))
nc2
TXx_d <- ncvar_get(nc2, 'TXx')
t <- ncvar_get(nc2, 'time')
ncatt_get(nc2, 'time')
summary(unlist(TXx_d))


# okay so so this has both of the TXx and the TXn 

TXx_d <- ncvar_get(nc2, 'TXx')
TXn_d <- ncvar_get(nc2, 'TXn')
```

```{r}
summary(unlist(TXx_d))
```

```{r}
summary(unlist(TXx_d) - unlist(TXn_d))
```


```{r}
t
```


```{r}
nc3 <- nc_open(file.path(INPUT, 'Rx1Day_CanESM2_historical_rcp85_r14i1p1_19500101-21001231.nc'))
Rx1day_DJF_d <- ncvar_get(nc3, 'Rx1day_MAM')
t <- ncvar_get(nc3, 'time')
ncatt_get(nc3, 'time')
summary(unlist(Rx1day_DJF_d))
```


```{r}
nc3
```


Okay so what I've learned here is that there are mulitple variables in each netcdf file, which is good because I was worried that some of the things were missing. This will make is difficult to designing the function but all the same is really good to know. 

So TN has both TNn and TNx 
TX has both TXn and TXx 

and then the percip values are broken up by different seasons and also an annual one which is awesome! 

```
Rx1day_DJF
Rx1day_MAM
Rx1day_SON
Rx1day_JJA
Rx1day_annual
```


# Exploring the data 

What would be easier, to do this in python, cdo, or R? 

```{r}
# Let's start with some of the temp data 
data <- TNx_d
t <- ncvar_get(nc1, 'time')
# Select the first lat and lon index at all of the different time steps
df <- data.frame(value = data[1, 1, ], 
                 time = t)

ggplot(data = df, aes(time, value)) + 
  geom_point() + 
  stat_smooth(method="lm") +
  stat_smooth(method="gam", color = 'red') +
  labs(title = 'Looking at Temp in 1 grid', y = 'K', x = 'year') ->
  x




```

```{r}

lm(data[1, 1, ] ~ t)
 
sub <- data[1:10, 1:10, ]
dim(sub) # so here is a our mini data 

```

# The most important part of the script! 

Here the functions for extracting the slope values are defined as well as the function that was used to create the functions 

```{r}
# The easiest thing to do is to flatten the dimensions of the files 
ntime <- dim(data)[3]
nlon  <- dim(data)[1]
nlat  <- dim(data)[2]

# Transfrom the data array that was read in from the netcdf file 
# into a 2D object that can is easier to apply functions to. 
# Args 
#   data: a 3D array (lon, lat, time)
#   nlon: dimensions of longitude
#   nlat: dimensions of latitude
# Returns: a 2D array (time, lon/lat)
reshape_3Darray <- function(data, nlon, nlat, ntime){
  
  # Make sure that we are reading in the 3d array 
  assert_that(length(dim(data)) == 3)
  
  # Parse out the lat/lon information. 
  nlon  <- dim(data)[1]
  nlat  <- dim(data)[2]
  ntime <- dim(data)[3]
  
  # Flatten out the array into a 2d array 
  flat      <- aperm(data, c(3, 2, 1))
  dim(flat) <- c(ntime, nlat * nlon)
  
  return(flat)
  
}

# Calculate change in trend via linear regression 
# Args 
#   data: a 2D array (time, lon/lat) created by the reshape_3Darray function 
#   nlon: dimensions of longitude
#   nlat: dimensions of latitude
# Returns: a matrix of the change by lon/lat
calculate_lm_slope <- function(data, nlon, nlat){
  
  # Make sure that the flat data is being read in 
  assert_that(length(dim(data)) == 2)
  
  # Now calculate the slope 
  slope_vector <- apply(data, 2, function(y){
    lm(y ~ t)$coefficients[2]
    })
  
  # Transform the results into a matrix by lat/lon 
  lm_results <- matrix(data = slope_vector, nrow = nlon, ncol = nlat, byrow = TRUE)

  return(lm_results)
}

# Calculate change via the difference in EOC
# Args
#   data: a 2D array (time, lon/lat) created by the reshape_3Darray function 
#   period_len: the length of the two different periods to take the average over  
#   nlon: dimensions of longitude
#   nlat: dimensions of latitude
# Returns: a matrix of the difference by lon/lat
calculate_EOC_diff <- function(data, nlon, nlat, period_len = 20){
  
  # Make sure that the flat data is being read in 
  assert_that(length(dim(data)) == 2)
  
  # Now calculate the difference between the historical and eoc mean. 
  diff_vector <- apply(data,2, function(y){
    
    # Historical average
    hist_avg <- mean(head(y, period_len))
    
    # The end of century average 
    eoc_avg <- mean(tail(y, period_len))
    
    # Calculate the difference 
    diff <- eoc_avg - hist_avg
    return(diff)
    
  })
  
  # Transform the results into a matrix by lat/lon 
  diff_results <- matrix(data = diff_vector, nrow = nlon, ncol = nlat, byrow = TRUE)
  
}

# DEFINE SOME FUNCTION THAT WILL DO IT THE THIRD WAY

# Make the netcdf 
# Args 
#   lm_rslts: the data returned by calculate_lm_slope
#   diff_rslts: the data returned by calculate_EOC_diff
#   lat: lattitude values 
#   lon: longitude values
#   info: a data frame containing information about the variable, model, senario,
#         and ensemble member. 
#   OUT_DIR: the location where to write the netcdf file out to 
make_nc <- function(lm_rslts, diff_rslts, lat, lon, info, OUT_DIR){
   
   # Check the inputs
   assert_that(dir.exists(OUT_DIR))
   assert_that(is.data.frame(info))
   assert_that(has_name(info, c('scn', 'model', 'ensemble', 'variable')))
   assert_that(length(dim(lm_rslts)) == 2)
   assert_that(length(dim(diff_rslts)) == 2)

   
   # Create the different dimensions for the netcdf file such as the time, lon, and lat
   dim_time <- ncdim_def('time', 'year', as.double(1)) # We know for a fact only 1 time step
   dim_lon  <- ncdim_def('lon', 'longitude', as.double(lon))
   dim_lat  <- ncdim_def('lat', 'latitude', as.double(lat))
   
   
   # Define the dimensions of the variables that will be used in the netcdf file. 
   lm_var   <- ncvar_def(name = "lm", units = 'NA', dim = list(dim_lon, dim_lat, dim_time),
                         longname = 'linear regression')
   diff_var <- ncvar_def(name = "diff", units = 'NA', dim = list(dim_lon, dim_lat, dim_time),
                         longname = 'difference between eoc and historical')
   
   # Name the netcdf file
   name    <- paste0(c(info$model, info$variable, info$scn, info$ensemble), collapse = '_')
   nc_name <- file.path(OUT_DIR, paste0(name, '.nc'))
   
   # Make the new netcdf file, but first delete the netcdf if it exists because 
   # the it is not possible to write over an already existing netcdf file. 
   if(file.exists(nc_name)){ file.remove(nc_name) }
   ncnew <- nc_create(nc_name, list(lm_var, diff_var), force_v4=TRUE)
   
   # Now that the ncdf is prepared, write the values to the netcdf. 
   ncvar_put(ncnew, "lm", vals = lm_rslts)
   ncvar_put(ncnew, "diff", vals = diff_rslts)
   
   # Define some global attributes 
   ncatt_put(ncnew, 0, "created_on", as.character(Sys.time()))
   ncatt_put(ncnew, 0, "scenario", info$scn)
   ncatt_put(ncnew, 0, "model", info$model)
   ncatt_put(ncnew, 0, "ensemble", info$ensemble)
   
   # Save the netcdf file. 
   nc_close(ncnew)
   return(nc_name)
   
 } 
 

```

```{r}
# data <- TNx_d
# ntime <- dim(data)[3]
# nlon  <- dim(data)[1]
# nlat  <- dim(data)[2]
# 
# d <- reshape_3Darray(data = data, nlon = nlon, nlat = nlat, ntime = ntime)
# lm_rslts   <- calculate_lm_slope(data = d, nlon = nlon, nlat = nlat)
# diff_rslts <- calculate_EOC_diff(data = d, nlon = nlon, nlat = nlat)
# 
# 
# make_nc(lm_rslts, diff_rslts, lat, lon, 
#         info = data.frame(model = 'this', scn = 'is', variable = 'bullshit', 
#                           ensemble = 'not', stringsAsFactors = FALSE),
#         OUT_DIR = '/Users/dorh012/Documents/2020/dev-NGD/tes_out')

```



